---
title: "oceancolour"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This package provides relatively easy access to [OceanColor](https://www.oceancolour.org/) through it's [OPeNDAP](https://www.oceancolour.org/thredds/catalog-cci.html) resources.

## Requirements

  + [R version 4.2+](https://www.r-project.org/)
  + [ncdf4](https://CRAN.R-project.org/package=ncdf4)
  + [stars](https://CRAN.R-project.org/package=stars)
  + [sf](https://CRAN.R-project.org/package=sf)

## Installation

Use the [remotes](https://CRAN.R-project.org/package=remotes) package from within R.
```
remotes::install_github("BigelowLab/oceancolour")
```

## Usage

The OceanColor products are quit large, so it is handy to be able to request and download a subset.  Here we use a bounding box for the Gulf of Maine.

We use a [URL](https://www.oceancolour.org/thredds/dodsC/cci/v6.0-1km-release/geographic/2024/ESACCI-OC-L3S-OC_PRODUCTS-MERGED-1D_DAILY_1km_GEO_PML_OCx_QAA-20240401-fv6.0_1km.nc.html) we have selected manually.  This is a high resolution (~ 1km) daily composite for just one day.


```{r}
suppressPackageStartupMessages({
  library(oceancolour)
  library(rnaturalearth)
  library(ncdf4)
  library(stars)
})

coast = rnaturalearth::ne_coastline(scale = "medium", returnclass = "sf")

url = "https://www.oceancolour.org/thredds/dodsC/cci/v6.0-1km-release/geographic/2024/ESACCI-OC-L3S-OC_PRODUCTS-MERGED-1D_DAILY_1km_GEO_PML_OCx_QAA-20240401-fv6.0_1km.nc"

bb = c(xmin = -72, xmax = -63, ymin = 39, ymax = 46)
var = "chlor_a"

NC1 = ncdf4::nc_open(url)
```

Check out `?get_lon` for a listing of helper functions used to navigate the object.  For example...

```{r}
(res1 <- get_res(NC1))
```

Now we generate a list that contains defines the bounding box, and us that to access the data. Not we request the output form [stars](https://r-spatial.github.io/stars/) which is the default.  We could also request "matrix".

```{r}
nav = get_nav(NC1, bb = bb)
x = get_var(NC1, var = var, nav = nav, form = 'stars')
x
```

And plot...

```{r}
plot(log10(x), axes = TRUE, reset = FALSE)
plot(sf::st_geometry(coast), add = TRUE, col = "darkgreen")
```


Finally, close the resource.

```{r}
ncdf4::nc_close(NC1)
```


### Aggregate data

Aggregate data files are also provided.  [This example](https://www.oceancolour.org/thredds/dodsC/CCI_ALL-v6.0-DAILY.html) provides an aggregate dataset of daily data for many variables from 1997 to present (lagged by a few months.)

```{r}
url = "https://www.oceancolour.org/thredds/dodsC/CCI_ALL-v6.0-DAILY"
NC2 = ncdf4::nc_open(url)
```

Note that the resource has much lower resolution.

```{r}
(res2 <- get_res(NC2))
```

Let's compare the high resolution example to this...

```{r}
res2/res1
```

So, the aggregate is about ~4km resolution.  But on the plus side it covers more time...

```{r}
summary(get_time(NC2))
```

... and ... contains a lot of variables...

```{r}
get_varnames(NC2)
```

Let's get the same date (2024-04-01) to compare.

```{r}
nav = get_nav(NC2, bb = bb, time = as.Date("2024-01-01"))
x2 = get_var(NC2, var = "chlor_a", nav = nav)
x2
```


```{r}
plot(log10(x2), axes = TRUE, reset = FALSE)
plot(sf::st_geometry(coast), add = TRUE, col = "darkgreen")
```

Finally, close this resource, too.

```{r}
ncdf4::nc_close(NC2)
```

